Causes and effects in Dichotomous Comparative Judgments: an information-theoretical system with plausible mechanism

### 2024-07-19 index.qmd
section:introduction; process:write

## Introduction {#sec-introduction}

@@

<!-- 2. in what has been applied on?, and where is the gap? -->
DCJ has proven effective in assessing competencies and traits predominantly within the educational realm, as demonstrated by @Pollitt_2012b, @Jones_2015, @vanDaal_et_al_2016, @Bartholomew_et_al_2018, @Lesterhuis_2018, @Bartholomew_et_al_2020, and @Marshall_et_al_2020. However, its application transcends education, as exemplified by @Boonen_et_al_2020. The methodology has also evolved to include multiple, as opposed to pairwise comparisons [@Luce_1959; @Placket_1975], and to accommodate comparisons with ordinal outcomes [@Tutz_1986; @Agresti_1992]. Overall, research suggests that DCJ offers an alternative and efficient approach to measurement and evaluation, characterized by its reliability and validity [@Lesterhuis_2018; @vanDaal_2020; @Marshall_et_al_2020]. Nevertheless, despite the method's widespread use, there is no clear representation in the literature of the plausible mechanisms that generate DCJ data. Particularly, there is no depiction of the complexity and the underlying assumptions of the DCJ system, nor how different assessment factors can potentially influence the observed DCJ outcome.

@@

<!-- 4. Solution to the gap -->
Although several studies have examined the individual impact of these factors on the method's reliability, including @Bramley_2015, @Pollitt_2012b, @Bramley_et_al_2019, @Verhavert_et_al_2019, @Crompvoets_et_al_2022, @vanDaal_et_al_2017, and @Gijsen_et_al_2021, to the best of the authors' knowledge, none have provided such transparent depiction of DCJ system and the plausible mechanisms that generate the DCJ outcome. This study aims to fill this gap by utilizing the framework of causal analysis and Directed Acyclic Graphs [DAG; @Pearl_2009; @Pearl_et_al_2016]. Using this framework, the study will construct a scientific model to elucidate the underlying assumptions of the DCJ system, providing plausible mechanisms of how the DCJ data is generated. This model will enable researchers to draw inferences about plausible causal relationships within the DCJ system. Furthermore, using a minimal set of assumptions from the framework, the study will translate the scientific model into a probabilistic statistical model, aiming to derive statistical estimands for different targets of inference. Ultimately, the results of this study could inform the planning of DCJ experiments and hold significance for researchers or analysts involved in education and assessment procedures who implement the DCJ methodology.

## Theoretical framework {#sec-TheoreticalFramework}

### Research questions and their estimands {#sec-ResearchQuestions}

### A scientific model for the DCJ procedure {#sec-SciMod}

### From the scientific to the Bradley-Terry-Luce model {#sec-Sci2StatMod}

## Discussion {#sec-discussion}

### Limitations and further research {#sec-LFR}

## Conclusion {#sec-conclusion}

@@

## Declarations {.appendix .unnumbered}

@@

## References {.appendix .unnumbered}

>>>>

# Introduction {#sec-introduction}

@@

<!-- 2. in what has been applied on?, and where is the gap? -->
DCJ has proven effective in assessing competencies and traits predominantly within the educational realm, as demonstrated by @Pollitt_2012b, @Jones_2015, @vanDaal_et_al_2016, @Bartholomew_et_al_2018, @Lesterhuis_2018, @Bartholomew_et_al_2020, and @Marshall_et_al_2020. However, its application transcends education, as exemplified by @Boonen_et_al_2020. The methodology has also evolved to include multiple, as opposed to pairwise comparisons [@Luce_1959; @Placket_1975], and to accommodate comparisons with ordinal outcomes [@Tutz_1986; @Agresti_1992]. Overall, research suggests that DCJ offers an alternative and efficient approach to measurement and evaluation, characterized by its reliability and validity [@Lesterhuis_2018; @vanDaal_2020; @Marshall_et_al_2020]. Nevertheless, despite the method's widespread use, the literature does not offer a clear representation of the plausible mechanisms that generate the DCJ data. Particularly, there is no depiction of the complexity and the underlying assumptions of the DCJ system, nor how different assessment factors can potentially influence the observed DCJ outcome.

@@

<!-- 4. Solution to the gap, final objective and relevance -->
Although several studies have examined the individual impact of these factors on the method's reliability [@Bramley_2015; @Pollitt_2012b; @Bramley_et_al_2019; @Verhavert_et_al_2019; @Crompvoets_et_al_2022; @vanDaal_et_al_2017; @Gijsen_et_al_2021], none, to the best of the authors' knowledge, have provided a transparent depiction of the DCJ system and the plausible mechanisms generating the DCJ outcome. This study aims to fill this gap by representing DCJ within the causal analysis framework [@Pearl_2009; @Pearl_et_al_2016]. Using Directed Acyclic Graphs [DAG, @Gross_et_al_2018; @Neal_2020], the study will construct a scientific model to elucidate the underlying assumptions of the DCJ system, providing plausible mechanisms for how the DCJ outcome could be generated. Next, using a minimal set of assumptions embedded in the framework, the study will translate the scientific model into a probabilistic statistical model. This model derives statistical estimands to draw inferences about plausible causal relationships within the DCJ system. Ultimately, this research aims to extend the law of comparative judgment initially proposed by @Thurstone_1927 and provide a sound probabilistic base for the statistical analysis of DCJ data. Consequently, this research holds significance for researchers and analysts involved in education and assessment procedures who implement or design DCJ experiments.

# Preliminaries {#sec-prelim}

## Causal analysis and DAGs {#sec-prelim-causal}

## The flow of association and causation {#sec-prelim-flow}

# Theoretical framework {#sec-theory}

## A scientific model for the DCJ {#sec-theory-scientific}

## From the scientific to statistical model {#sec-theory-statistics}

## Where does it all fit? {#sec-theory-where}

# Discussion {#sec-discuss}

## Findings {#sec-discuss-finding}

## Limitations and further research {#sec-discuss-limitations}

# Conclusion {#sec-conclusion}

@@

# Declarations {.appendix .unnumbered}

@@

# Appendix {#sec-appendix}

## Why estimate judges appart? {#sec-appA}

<!-- The same idea as interactions: when interactions are significant, the interpretations of the parameters cannot be made in isolation -->

## References {.unnumbered}
